---
title: "Exam 3"
output: 
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: true
      smooth_scroll: true
---
Adam Hales

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Setup
```{r}
library(tidyverse)
library(broom)
```

---

## Task 1

### Read in csv
```{r}
#Read in csv 
dat <- read.csv("FacultySalaries_1995.csv")
```


### Tidy data for plotting
```{r}
dat <- dat %>% 
  filter(Tier %in% c("I","IIA", "IIB")) %>% 
  select("Tier", "State",
         Full = "AvgFullProfSalary", 
         Assoc = "AvgAssocProfSalary", 
         Assist = "AvgAssistProfSalary") %>% 
  pivot_longer(cols = c("Full", "Assoc", "Assist"),
               names_to = "Rank",
               values_to = "Salary")
```

### Creating the Plot
We want to make a boxplot showing the difference in salary for professors at different tiers of institutions. 
```{r}
p1_dat<- 
  ggplot(dat)+
  aes(x = Rank, y = Salary, fill = Rank)+
  geom_boxplot()+
  facet_wrap(~Tier)+
  labs( 
    x = "Rank", 
    y = "Salary", 
    fill = "Rank" 
  ) + 
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45)
  ) 

print(p1_dat)
```

---

## Task 2
Given that cleaned data set, we will build an ANOVA model and display the summary. <br>
This model will test the influence of "State" "Tier" and "Rank" on "Salary" but not interactions between predictors.

```{r}
#verify factors are set before making model
head(dat)
#set factor levels
dat$Tier <- factor(dat$Tier)
dat$Rank <- factor(dat$Rank)

#make model
anova_mod <- aov(Salary ~ State + Tier + Rank, data = dat)

#get model summary
summary(anova_mod)
```
As you can see, state, tier, and rank are statistically significant predictors of professor salary. 

---

## Task 3
Now we will switch data to examine the juniper oils data set.
```{r}
#read in csv
juniper <- read.csv("Juniper_Oils.csv")
```

### Select Necessary Info
This data set has a bunch of stuff that we don't need. Let's clean it up. <br>
We really only need the chemicals and years since burn. 
```{r}
important_juniper <- juniper %>% 
  select(YearsSinceBurn, alpha.pinene:thujopsenal)
```

### Tidy Data
```{r}
long_juniper <- important_juniper %>% 
  pivot_longer(cols = alpha.pinene:thujopsenal,
               names_to = "Chemical_name",
               values_to = "Concentration")
```

## Task 4
Make a plot showing the following:
   
- x = YearsSinceBurn
- y = Concentration
- facet = ChemicalID (use free y-axis scales)

```{r}
ggplot(long_juniper)+
  aes(x = YearsSinceBurn, y = Concentration)+
  facet_wrap(~Chemical_name, scales = "free_y")+
  geom_smooth(color = "blue")+
  theme_minimal() +
  theme(
    strip.text = element_text(size = 7) #make text just a little smaller so you can see the full chemical name
  ) +
  labs(
    title = "Chemical Concentrations and Years Since Burn",
    x = "Years Since Burn",
    y = "Concentration"
  )
```

---

## Task 5
Use a generalized linear model (glm) to find which chemicals show concentrations that are significantly affected by years since burn.
__Assume significance = p<0.05.__

To do this, I chose to make one large model and then narrow it down based on significant predictors.

```{r}
#Make one big model and then we can clean it up later
juniper_mod_big <- glm(Concentration ~ YearsSinceBurn * Chemical_name, 
                       data = long_juniper)

#tidy model output
mod_output <- tidy(juniper_mod_big)

#filter for significant interactions
significant_chemicals <- mod_output %>% 
  filter(p.value < 0.05) #filter for significance

print(significant_chemicals)
```

